---
title: "Week 7 - Unsupervised Clustering for Segmentation"
author: "Chris Healey"
output:
  html_document:
    toc: true
    theme: united
---

I want to use a common ML dataset, Titanic passengers, for our examples this week.

Our main goal is to identify characteristics of passenger survival using unsupervised clustering.

- introducing k-means clustering methods and issues    
- interpretation of these result of this clustering

We'll take a look at supervised methods in another notebook.

# k-means clustering
In this section, we try to identify clusters based on K-Means clustering.  

## Data setup
It's important to note that K-Means works on real-valued and non-binary variables.  Additionally, k-means clustering tends to work a little better on normalized variables.  As long as every column is numeric, we can normalize each column using the `scale` function.

```{r}
library(tidyverse)

#load data
df <- read_csv("data/titanic.csv")
```


```{r}
#encode the variable for sex
df$Female = ifelse(df$Sex=="female",1,0)

#create new dataframe with only numeric columns
#drop character variables
points <- df %>% select(-c("Name","Sex"))
print(head(points))
```


```{r}
points <- scale(points)
print(head(points))
```

## k-means Basics
We have to specify the number of clusters ahead of time, but the method is relatively easy to run.
```{r}
#set seed for code reproducibility
set.seed(1223)
NUM_CLUSTERS <- 6
kclust <- kmeans(points, centers = NUM_CLUSTERS)

#add segments to original dataset
df$kmeans_segment <- as.factor(kclust$cluster)

print(df %>% relocate(kmeans_segment))
```

## Choosing a good number of clusters
We may want to investigate how the total within cluster variation changes as we increase the number of segments.
```{r}
ks <- 1:12
tot_within_ss <- sapply(ks, function(k) {
    set.seed(1223)
    cl <- kmeans(points, k)
    cl$tot.withinss
})
plot(ks, tot_within_ss, type = "b")
```
Remember we want be able to keep the number of clusters as small as possible while still providing insights!

# Interpretations and Visualizations

Based on the graph above, I'm going to set the number of clusters equal to 5 and 6 ... and something interesting and illustrative of cluster analysis is going to happen....
```{r}
set.seed(1223)
NUM_CLUSTERS <- 5
kclust <- kmeans(points, centers = NUM_CLUSTERS, nstart=10)

#add segments to original dataset
df$kmeans_segment <- as.factor(kclust$cluster)

```

## Segment profiles
For the segment profiles, I summarize the results of each segment (using the original dataset columns -- without the scaling).

```{r}
df %>% select(-c("Name","Sex")) %>% add_count(kmeans_segment) %>%
  group_by(kmeans_segment,n) %>% 
  summarise_all("mean")
```

Altogether, we do this in one piece of code
```{r}
set.seed(1223)
NUM_CLUSTERS <- 6
points <- df %>% select(-c("Name","Sex","kmeans_segment"))
points <- scale(points)
kclust <- kmeans(points, centers = NUM_CLUSTERS, nstart = 10)

#add segments to original dataset
df$kmeans_segment <- as.factor(kclust$cluster)

df %>% select(-c("Name","Sex")) %>% add_count(kmeans_segment) %>%
  group_by(kmeans_segment,n) %>% 
  summarise_all("mean")

```
## Example 2-D Visualization
Here's visualization of the segments featuring two real-valued variables.
```{r}
g <- ggplot(df)+geom_point(aes(x=Fare,y=Age,colour=kmeans_segment)) +scale_x_continuous(trans='log10')
print(g)
```
This picture is a little muddy because we didn't train the model on these variables (we trained it on all columns)







## Another approach
Just because we have all columns available, we don't necessarily have to use them.  For example, maybe we want to investigate `fare` and `age` alone.
```{r}
set.seed(1223)
points <- df %>% select(Age,Fare)
points <- scale(points)
kclust <- kmeans(points, centers = NUM_CLUSTERS, nstart = 10)

#add segments to original dataset
df$kmeans_segment <- as.factor(kclust$cluster)

df %>% select(-c("Name","Sex")) %>% 
  group_by(kmeans_segment) %>% 
  summarise_all("mean")
```


```{r}
g <- ggplot(df)+geom_point(aes(x=Fare,y=Age,colour=kmeans_segment))+scale_x_continuous(trans='log10')
print(g)

```

